From Reversible Programs to \\ Univalent Universes and Back

### 2021-09-05 Pi+/paper/finite.tex
Fix part of the typos

\begin{proposition}
  \leavevmode
  \begin{enumerate}
    \item $\normf$ splits into a section-retraction pair, that is, we have ${\List[\Fin[n]]} \xrightarrow{s} \Sn[n]
            \xrightarrow{r} {\List[\Fin[n]]}$ such that $s \comp r \htpy \normf$ and $r \comp s \htpy \idfunc_{\Sn[n]}$.
    \item \(\im{\quotinc} \eqv \Sn \eqv \im{\normf} \), where $\quotinc$ is the inclusion into the quotient, and
          $\im{f}$ denotes the image of $f$ (see~\cref{def:im}).
  \end{enumerate}
\end{proposition}

@@

\begin{definition}[Inversion count]
  Given a permutation $\sigma : \Aut[\Fin[n]]$, the inversion count of $k: \Fin[n]$ is given by
  \[ \invcount{i} = \#\Set{j < i | \sigma(j) > \sigma(i)}. \]
\end{definition}

@@

\noindent To decode the permutation back from this Lehmer code, we perform an algorithm similar to \emph{insertion
  sort}. The element of the Lehmer code being currently processed is highlighted in the left column of the table below.
Starting from a sorted list, the element at index $k$ has to be given $l[k]$ inversions. Because of the the invariant
that all the elements before newly processed one are smaller than it, the proper number of inversions is created by
simply shifting the element $l[k]$ places left.

@@

\begin{theoremrep}
  \leavevmode
  \begin{enumerate}
    \item For any Lehmer code $c$, $\immersion_{n}(c)$ is a normal form with resepct to $\longcoxstar$, that is,
          $\immersion_{n}(c)$ is in $\im{\normf}$.
    \item Any element of $\im{\normf}$ can be constructed from a unique Lehmer code by $\immersion$, that is, the fibers
          of $\immersion_{n} : \Lehmer[n] \to {\im{\normf}}$ are contractible.
  \end{enumerate}
  Therefore, there is an equivalence between $\Lehmer[n]$ and $\im{\normf}$.
\end{theoremrep}

>>>>

\begin{proposition}
  \leavevmode
  \begin{enumerate}
    \item $\normf$ splits into a section-retraction pair, that is, we have ${\List[\Fin[n]]} \xrightarrow{s} \Sn[n]
            \xrightarrow{r} {\List[\Fin[n]]}$ such that $s \comp r \htpy \normf$ and $r \comp s \htpy \idfunc_{\Sn[n]}$.
    \item \(\im{\quotinc} \eqv \Sn \eqv \im{\normf} \), where $\quotinc$ is the mapping into the quotient, and
          $\im{f}$ denotes the image of $f$ (see~\cref{def:im}).
  \end{enumerate}
\end{proposition}

@@

\begin{definition}[Inversion count]
  Given a permutation $\sigma : \Aut[\Fin[n]]$, the inversion count of $i: \Fin[n]$ is given by
  \[ \invcount{i} = \#\Set{j < i | \sigma(j) > \sigma(i)}. \]
\end{definition}

@@

\noindent To decode the permutation back from this Lehmer code, we perform an algorithm similar to \emph{insertion
  sort}. The element of the Lehmer code being currently processed is highlighted in the left column of the table below.
Starting from a sorted list, the element at index $k$ has to be given $l[k]$ inversions. Because of the invariant
that all the elements before newly processed one are smaller than it, the proper number of inversions is created by
simply shifting the element $l[k]$ places left.

@@

\begin{theoremrep}
  \leavevmode
  \begin{enumerate}
    \item For any Lehmer code $c$, $\immersion_{n}(c)$ is a normal form with respect to $\longcoxstar$, that is,
          $\immersion_{n}(c)$ is in $\im{\normf}$.
    \item Any element of $\im{\normf}$ can be constructed from a unique Lehmer code by $\immersion$, that is, the fibers
          of $\immersion_{n} : \Lehmer[n] \to {\im{\normf}}$ are contractible.
  \end{enumerate}
  Therefore, there is an equivalence between $\Lehmer[n]$ and $\im{\normf}$.
\end{theoremrep}

### 2021-09-05 Pi+/paper/ufin.tex
Fix part of the typos

Functions between types are functors between groupoids, and type families (or functions to the universe) are indexed
families of groupoids. A type family $P : A \to \UU$ comes equipped with a functor $\pi_1 : \dsum*{x:A}{P(x)} \to \UU$
which has a lifting operation giving it the structure of a fibration. The $\term{transport}$ operation lifts paths in
the base space to functions between the fibers.  Using the groupoid structure of $A$, for a path $p : x \id_{A} y$,
$\term{transport}(p)$ and $\term{transport}(\inv{p})$ form an equivalence.
% We name this map $\tptEqv{P}$ which lifts paths to equivalences of fibers.
\[
  \tptEqv{P} : \dfun*{x,y:A}{(x \id_{A} y) \to (P(x) \eqv P(y))}
\]

@@

This definition of the groupoid of finite types has also been considered
in~\cite{yorgeyCombinatorialSpeciesLabelled2014}. While $\UFin$ has \emph{all} the finite types, we are also interested
in constructing a subuniverse of finite types of a specified cardinality. To do so, we will start with the subuniverse
$\BAut[T]$. \footnote{Characterisations of univalent fibrations using the $\BAut$ construction have been studied
  by~\citet{christensenCharacterizationUnivalentFibrations2015}.}

@@

Similar to $\BFin$, the groupoid $\UFin$ has two symmetric monoidal structures, the additive and the multiplicative one,
and the multiplicative tensor product distributes over the additive one. To construct these, we first state and prove
some equivalences on $\Fin$, and some general type isomorphisms. Then we simply lift these equivalences to $\UFin$, by
the univalence principle.

\begin{proposition}
  For any $n, m : \Nat$, and for any types $X, Y, Z$,
  \begin{gather*}
    \begin{aligned}
      \Fin[0]                & \eqv \bot                  \\
      \Fin[n] \sqcup \Fin[m] & \eqv \Fin[n + m]           \\
      \\
      \bot \sqcup X          & \eqv X                     \\
      X \sqcup \bot          & \eqv X                     \\
      (X \sqcup Y) \sqcup Z  & \eqv X \sqcup (Y \sqcup Z) \\
      X \sqcup Y             & \eqv Y \sqcup Y            \\
      X \times \bot          & \eqv \bot
    \end{aligned}
    \qquad
    \begin{aligned}
      \Fin[1]                & \eqv \top                             \\
      \Fin[n] \times \Fin[m] & \eqv \Fin[n * m]                      \\
      \\
      \top \times X          & \eqv X                                \\
      X \times \top          & \eqv X                                \\
      (X \times Y) \times Z  & \eqv X \times (Y \times Z)            \\
      X \times Y             & \eqv Y \times X                       \\
      X \times (Y \sqcup Z)  & \eqv (X \times Y) \sqcup (X \times Z)
    \end{aligned}
  \end{gather*}
\end{proposition}

>>>>

Functions between types are functors between groupoids, and type families (or functions to the universe) are indexed
families of groupoids. A type family $P : A \to \UU$ comes equipped with a functor $\pi_1 : \dsum*{x:A}{P(x)} \to A$
which has a lifting operation giving it the structure of a fibration. The $\term{transport}$ operation lifts paths in
the base space to functions between the fibers.  Using the groupoid structure of $A$, for a path $p : x \id_{A} y$,
$\term{transport}(p)$ and $\term{transport}(\inv{p})$ form an equivalence.
% We name this map $\tptEqv{P}$ which lifts paths to equivalences of fibers.
\[
  \tptEqv{P} : \dfun*{x,y:A}{(x \id_{A} y) \to (P(x) \eqv P(y))}
\]

@@

This definition of the groupoid of finite types has also been considered
in~\cite{yorgeyCombinatorialSpeciesLabelled2014}. While $\UFin$ has \emph{all} the finite types, we are also interested
in constructing a subuniverse of finite types of a specified cardinality. To do so, we will start with the subuniverse
$\BAut[T]$, for any type $T : \UU$. \footnote{Characterisations of univalent fibrations using the $\BAut$ construction
have been studied by~\citet{christensenCharacterizationUnivalentFibrations2015}.}

@@

Similar to $\BFin$, the groupoid $\UFin$ has two symmetric monoidal structures, the additive and the multiplicative ones,
and the multiplicative tensor product distributes over the additive one. To construct these, we first state and prove
some equivalences on $\Fin$, and some general type isomorphisms. Then we simply lift these equivalences to $\UFin$, by
the univalence principle.

\begin{proposition}
  For any $n, m : \Nat$, and for any types $X, Y, Z$,
  \begin{gather*}
    \begin{aligned}
      \Fin[0]                & \eqv \bot                  \\
      \Fin[n] \sqcup \Fin[m] & \eqv \Fin[n + m]           \\
      \\
      \bot \sqcup X          & \eqv X                     \\
      X \sqcup \bot          & \eqv X                     \\
      (X \sqcup Y) \sqcup Z  & \eqv X \sqcup (Y \sqcup Z) \\
      X \sqcup Y             & \eqv Y \sqcup X            \\
      X \times \bot          & \eqv \bot
    \end{aligned}
    \qquad
    \begin{aligned}
      \Fin[1]                & \eqv \top                             \\
      \Fin[n] \times \Fin[m] & \eqv \Fin[n * m]                      \\
      \\
      \top \times X          & \eqv X                                \\
      X \times \top          & \eqv X                                \\
      (X \times Y) \times Z  & \eqv X \times (Y \times Z)            \\
      X \times Y             & \eqv Y \times X                       \\
      X \times (Y \sqcup Z)  & \eqv (X \times Y) \sqcup (X \times Z)
    \end{aligned}
  \end{gather*}
\end{proposition}
